The \project{unRAVE} catalog

### 2016-09-08 article/unrave.tex
Implement MKN suggestion. Closes #27

\begin{itemize}
\item We assume that any fibre- and time-dependent variations in spectral
resolution in the \rave\ spectra are negligible.
\item The \rave\ noise variances are approximately correct, independent between
pixels, and normally distributed.
\item We assume that the normalization procedure employed by the \rave\ pipeline
is invariant with respect to the labels we seek to measure (e.g., $\teff$, $\logg$,
or [Fe/H]), and invariant with respect to the S/N ratio.  In other words, we assume
that the normalization procedure does not produce different results for high S/N
spectra compared to low S/N spectra, nor does the normalization procedure vary 
non-linearly with respect to stellar parameters (e.g., [Fe/H]).
\item We assume that stars with similar labels ($\teff$, $\logg$, and abundances)
have similar spectra.
\item A stellar spectrum is a smooth function of the label values for that star,
and we assume that the function is smooth enough within a sub-space of the labels
(e.g., the giant branch or the main-sequence) that it can be reasonably approximated 
with a low-order polynomial in label space.
\item The training set (Section~\ref{sec:the-training-set}) has mean accurate labels.
Note that we do not assume that \emph{every} label in the training is accurate -- we
can afford to have a small fraction of inaccurate labels (e.g., a few obvious 
misclassifications in the training set are affordable).
\item We assume that the training data are similar (in spectra) to the test data 
where they overlap in label space, and we assume that the training data spans enough
of the label space to capture the variation in the test-set spectra.
\end{itemize}

@@

We constructed a three-label quadratic model using only main-sequence and sub-giant
stars. In order to set the regularization hyperparameter
$\Lambda$ for this model, we trained 30 models with different regularization strengths,
spaced evenly in logarithmic steps between $\Lambda = 10^{-3}$ to $\Lambda = 10^{3}$.
We then performed leave-one-out cross-validation for each model.  Specifically, for 
each star in the training set: we removed the star; trained the model; and then 
inferred labels from the removed star as if it was a test object. We also performed 
leave-one-out cross-validation on an unregularized ($\Lambda = 0$) model, which we 
will use as the basis for comparison.  For the unregularized case, we calculated the 
bias and root-mean-square (RMS) deviation between: the training set labels, and the 
labels we derive by cross-validation, where one star is removed at a time and the
model is re-trained.  We repeated this calculation of bias and RMS deviation for all
30 models with different regularization strengths $\Lambda$.

We show the \emph{percentage difference} in the RMS deviation of the labels with respect
to the unregularized model in Figure \ref{fig:set-hyperparameters}.  The upper 
and lower envelope represent the boundaries across all labels, showing that with increasing
regularization, the RMS decreased in \emph{all} labels.  We found similar
improvements in the biases, however these were already minimal in the unregularized
case.  The improvement in RMS reaches a minimum value near 
$\Lambda = 35.6$ ($\approx10^{1.5}$), where
we achieve RMS deviations that are about 10\% better than the unregularized case.
Based on this improvement we set $\Lambda = 35.6$ for this model.  At this regularization 
strength, the bias and RMS values found by leave-one-out
cross validation are, respectively: $38$~K and $256$~K for $\teff$, $0.05$~dex and 
$0.29$~dex for $\logg$, with $0.03$~dex and $0.17$~dex for [Fe/H].

We inferred labels for all \rave\ spectra using this main-sequence/sub-giant star model.  
The results for the survey sample are shown in the centre panel of Figure \ref{fig:test-set-density}.  
The increased density of solar-type stars is consistent with \rave\
observing stars in the local neighbourhood, and the high number of turn-off
and main-sequence stars relative to the sub-giant branch is expected from 
the relative lifetimes of these evolutionary phases.  An over-density
of stars near the base of the giant branch is also present.  This artefact is due 
to having giant stars in the test set, but not in the training set, and the
model is (poorly) extrapolating outside the convex hull of the training set.

\subsection{A 9-label model for detailed abundances of giant stars}
\label{sec:evolved-star-model}

The red giant branch stars in our training set have stellar parameters 
($\teff$, $\logg$) and up to 15 elemental abundances from the \aspcap\ 
\citep{Garcia_Perez_2016}.  A subset of these elements have atomic transitions in the 
\rave\ wavelength region: \ion{O}{1}, \ion{Mg}{1}, \ion{Al}{1}, \ion{Si}{1}, 
\ion{Ca}{2}, \ion{Ti}{1}, \ion{Fe}{1}, and \ion{Ni}{1}.  However, we 
excluded [Ti/H] from our abundance list because of systematics in the \aspcap\
[Ti/H] abundances \citep{Holtzman_2015,Hawkins_2016}.  Therefore we are left 
with nine labels in our giant star model: $\teff$, $\logg$, and seven elemental 
abundances.

Similar to Sections \ref{sec:a-simple-model} and \ref{sec:unevolved-star-model},
we used a quadratic vectorizer for the giant star model.  Here the terms are 
expanded in the same way as equation \ref{eq:vectorizer-three-label}, only
with nine labels instead of three.  We set the regularization hyperparameter $\Lambda$
in the same way described in Section \ref{sec:unevolved-star-model}, using the same 30 trials of $\Lambda$.
The results are shown in Figure \ref{fig:set-hyperparameters}, where
again the enveloped region represents the minimum and maximum change in RMS label
deviation with respect to the unregularized case.  At the point of maximum improvement 
near $\Lambda = 0.13$ $(\approx10^{-0.9})$, the RMS in all nine labels has decreased by up to $30$\%,
with all labels showing an improvement $>5$\%, and the mean improvement over all labels is about 10\%.
Near $\Lambda \approx 10^{-0.9}$ to $10^{-0.3}$, the regularization also produces a sparser matrix
of $\vectheta$, with $\approx20$\% more terms (mostly cross-terms) having zero-valued entries.
Based on the increased model sparsity and decreasing RMS deviation in the labels, 
we adopt $\Lambda = 0.57$ ($10^{-0.25}$) for the giant star model.  The bias in 
labels from a regularized model with $\Lambda = 0.57$ is negligible: $-0.3$~K in 
$\teff$, and $<$0.007~dex in magnitude for $\logg$ and all seven elements.  The RMS
at this regularization strength is 69~K in $\teff$, 0.18~dex in $\logg$, and varies
between $0.07-0.09$~dex depending on the elemental abundance.  We inferred labels 
for all \Nspectra\ \rave\ spectra using this model, and these results are summarized in
the right panel of Figure \ref{fig:test-set-density}.  The red clump is clearly visible
and in the expected location, without requiring any post-analysis calibration.
However, artefacts due to dwarf stars in the test set are also present.

\subsection{Joining the models}
\label{sec:joining-the-models}

We have derived labels for all \Nspectra\ \rave\ spectra using the three models
described in previous sections.  The results from the joint model --- that includes 
the main-sequence, sub-giant and red giant branch --- shows that a single 3-label quadratic
model is too simple for the \rave\ spectral range.  The other models have problems,
too: unrealistic over-densities in label space show that the main-sequence model and
the giant model make very poor extrapolations for stars outside their respective
training sets.  For these reasons we were forced to exclude or severely penalize 
incorrect results from both models.  We emphasize that the choices here are entirely
heuristic, and depart from interpreting \thecannon\ output labels as the maxima
of individual likelihood functions.  Each model produces estimates of the labels
for a given star, and we use those estimates to produce a unified estimate, but 
this joint estimate is calculated by disregarding the probabilistic attributes
of individual estimates.

Before attempting to join the results from different models, we excluded results
in either model that had a reduced $\chi_{r}^2 > 3$.  We further discarded stars with
labels that are outside the extent of the training set.  Specifically for the
results from the giant model we (conservatively) excluded stars with derived 
$\logg > 3.5$, and for the results from the main-sequence model we excluded 
sub-giant stars ($\logg < 4$ and $\teff < 5000$~K) that were outside the 
two-dimensional ($\teff$, $\logg$) convex hull of the training set used for the main-sequence model.  Unfortunately these restrictions did not remove all spurious
results.  The reason for this can be explained with an example:  consider that our 
giant star model was trained with only giant stars but tested with both giant 
stars and dwarf stars.  Some classes of stars (e.g., metal-poor dwarfs) can 
project into a region of label space that would suggest it is a giant 
(e.g., a clump star).  These objects could have relatively low $\chi_{r}^2$ values
(e.g., $\chi_{r}^2 < 3$) and in this example, they would appear as bonafide red clump 
stars.  These incorrect projections are extrapolation errors in high dimensions that
project to `normal' parts of the label space in two dimensions.  For these reasons 
we also made use of the joint model in Section~\ref{sec:a-simple-model}
to inform whether we should adopt results from: the red giant branch model; the
main-sequence/sub-giant model; or a linear combination of the two.

In Figure \ref{fig:joint-model-differences} we show the differences in effective temperature
$\teff$ and surface gravity $\logg$ between: the main-sequence model and the joint
model; and the differences between the red giant branch model and the joint model. 
We have scaled the differences in $\teff$ and $\logg$ to make the central peak 
near $(0, 0)$ to be approximately isotropic by setting
	$\delta_{T_{\rm eff}} = 90$~K for the main-sequence model, 
	$\delta_{T_{\rm eff}} = 50$~K for the giant model, and 
	$\delta_{\log{g}} = 0.15$~dex for both models.
It is important to note that these $\delta$ values do not represent any kind of intrinsic
uncertainty or precision: they are merely normalization factors.  Adopting
substantially different scaling factors would produce in clear inconsistencies in our
results (e.g., sub-giants being misclassified as giants).  Therefore we chose these 
factors empirically to make the distributions in Figure \ref{fig:joint-model-differences}
approximately isotropic, and to some extent, comparable.  In Figure 
\ref{fig:joint-model-differences} the stars within the peak at $(0, 0)$ represent objects where
the joint model and the comparison model both report similar labels.  
The artefacts 
seen in the Hertzsprung-Russell diagrams in Figure \ref{fig:test-set-density} are
also present in Figure \ref{fig:joint-model-differences} as over-densities far away
from the central peak.  Therefore, we can 
adopt the scaled distance in labels $\teff$ and $\logg$ from the joint model to the 
main-sequence model $d_{ms}$,
\begin{eqnarray}
	d_{ms} & = & \left(\frac{T_{{\rm eff},ms} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},ms}}}\right)^2 + \left(\frac{\log{g}_{ms} - \log{g}_{joint}}{\delta_{\log{g},ms}}\right)^2 \quad .
\end{eqnarray}

However when adopting a similar distance metric for the giant model, we found that the
giant model and the joint model produced similar label estimates (with low $\chi_r^2$
values) for metal-poor ($[{\rm Fe/H]} < -1$) dwarf stars near the turn-off.  The similarity
in estimated labels occurred in $\teff$, $\logg$, \emph{and} [Fe/H], implying that we
could not use the difference between $[{\rm Fe/H}]_{giant} - [{\rm Fe/H]}]_{joint}$
as a term that would benefit the main-sequence model.  As a result, our joint estimate 
for metal-poor dwarf stars that overlap with high-resolution studies
\citep{Reddy_2003,Reddy_2006,Bensby_2014,Valenti_Fischer_2005},
would appear as metal-poor giant stars, because the giant star model would have a much
smaller distance metric than the main-sequence model.  It is reasonable to assert that
this issue may be complicated by the increased complexity, or flexibility, of the giant
star model relative to the main-sequence model.  In any case, we found that by empirically
penalizing the giant model distance metric by absolute metallicity,
\begin{eqnarray}
	d_{giant} & = & \left(\frac{T_{{\rm eff},giant} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},giant}}}\right)^2 + \left(\frac{\log{g}_{giant} - \log{g}_{joint}}{\delta_{\log{g},giant}}\right)^2  + \lvert[{\rm Fe/H}]_{joint}\rvert  \quad ,
\end{eqnarray}

\noindent{}the weighting of dwarfs was no longer a function of joint metallicity, and this
issue was resolved without introducing new artefacts. Using these distance metrics we 
then derived the weights,
\begin{eqnarray}
	w_{ms} = \frac{1}{{d_{ms}}^2} & \text{and} & w_{giant} = \frac{1}{{d_{giant}}^2} \quad ,
\end{eqnarray}

\noindent{}and produce the weighted labels $\hat\ell$:
\begin{eqnarray}
	\hat\ell = \frac{w_{ms}\,\ell_{ms} + w_{giant}\,\ell_{giant}}{w_{ms} + w_{giant}} \quad .
\end{eqnarray}

We calculate weighted errors of $\hat\ell$ in the same manner.  
In Figure \ref{fig:model-weights} we show the mean relative
weight $w_{ms}/(w_{ms} + w_{giant})$ within each two-dimensional bin of 
$\hat\teff$ and $\hat\logg$.  For giant stars the relative weight of 
the main-sequence model is zero, and vice-versa for main-sequence stars.
The relative weights smoothly transition from 0 to 1 on the sub-giant branch
near $\log{g} \approx 3.5$, in the training set overlap region of both models.
For abundance labels in the giant model that are not in the main-sequence
model (e.g., [O/H], [Mg/H]), we only report abundances for objects if
$w_{ms} < 0.05$.

The weighted $\teff$ and $\logg$ values for stars meeting different
S/N constraints are shown in Figure \ref{fig:test-set-hrd}, both in logarithmic density
and mean metallicity.  The artefacts from individual models are no longer 
apparent, and the rich structure of the Hertzsprung-Russell is visible.  Despite
this, there are a number of caveats introduced by the choices we have made to
combine estimates from multiple models. We discuss these issues in detail in
Section \ref{sec:discussion}.

\section{Validation experiments}
\label{sec:validation}

In addition to the cross-validation tests that we have previously described, 
we have conducted a number of internal and external validation experiments to 
test the validity of our results.  We will begin by describing internal validation
tests based on repeat observations, before evaluating our accuracy based on
high-resolution literature comparisons.

\subsection{Multi-epoch observations}
\label{sec:repeat-observations}

The \rave\ survey performed repeat observations for 43,918 stars with time 
intervals ranging from a few hours to up to four years.  This timing was 
constructed to be quasi-logarithmic such that spectroscopic binaries could
be optimally identified. Most of the stars that were observed multiple times
were only observed twice, with thirteen visits being the maximum number 
of observations for any target.  These repeat observations allow us to 
quantify the level of (in)correctness in our formal errors.  For every star
with multiple visits we constructed a high S/N stacked spectrum for that
star, weighted by the inverse variances in the individual visit spectra.

We treated the stacked spectra as normal survey stars.  We inferred labels
using all models and joined the results as per Section \ref{sec:joining-the-models}.  The labels inferred from the stacked spectra served as the basis
of comparison for all individual visits to that object.  Figure 
\ref{fig:formal-errors-comparison} shows the difference in labels 
between the comparison spectrum and a repeat visit, normalized by their 
formal errors summed in quadrature (e.g., 
$\Delta\logg/\sqrt{\sigma_{\logg,stacked}^2 + \sigma_{\logg,visit}^2}$).
If our measurements were unbiased by S/N and the formal errors were 
representative, these values should be normally distributed with a zero 
mean and unit variance.
\stub{Are they? (Probably not!). How much should we inflate our formal errors, and should this be as a function of S/N? Note that in all Figures and tables,
we show these inflated errors only, as we gauge them to be more representative.}

\stub{In Figure \ref{fig:repeat-visits} we look at the difference as a function of SNR}

\subsection{External validation}
\label{sec:external-validation}

\subsubsection{Comparison with \rave\ DR4}
\label{sec:validation-kordopatis}

As an initial point of comparison, we cross-matched our results against the 
official fourth \rave\ data release (Figure \ref{fig:rave-dr4-comparison}).
In order to provide a fair comparison, we only show stars that meet a number
of quality flags in \emph{both} samples.  Our constraints require that the
S/N ratio exceeds 10~pixel$^{-1}$, and $\chi_r^2 < 3$.  For this comparison
we further required that:
the \texttt{QK} flag from \citet{Kordopatis_2013} is zero, indicating no
problems were reported by the pipeline; $T_{{\rm eff},DR4} > 4000$~K;
the error in radial velocity \texttt{e\_HRV} is $<$8~km~s$^{-1}$; and the three
principal morphological flags \texttt{c1}, \texttt{c2}, \texttt{c3}, from 
\citet{Matijevic_2012} all indicate `n' for a normal FGK-type star.
There is good agreement in $\teff$, with a bias and RMS of just 4~K and 240~K,
respectively. The offset in $\logg$ on the giant branch between this study and 
\citet{Kordopatis_2013} has been noted in other studies (e.g., \apogee), and this issue
has been resolved in the fifth \rave\ data release by correcting $\logg$
values with a calibration sample consisting of asteroseismic targets and the
\gaia\ benchmark stars.  There is also a
slight discrepancy in the $\logg$ values along the main-sequence, where our
work tends to taper down towards higher $\logg$ values at cooler temperatures,
and the \rave\ DR4 sample tends to have a slightly flatter lower main-sequence.
This difference is not likely to have a very significant effect on the detailed
abundance or spectrophotometric distance determinations between these studies 
\citep{Binney_2014}.

\subsubsection{Comparisons with Reddy, Bensby, and Valenti \& Fischer}
\label{sec:validation-gold-standards}

We searched the literature for studies that overlap with \rave, and which base
their analysis on high-resolution, high S/N spectra.  We found four notable
studies with a sufficient level of overlap: the disk studies by \citet{Reddy_2003,Reddy_2006}
and \citet{Bensby_2014}, as well as the \citet{Valenti_Fischer_2005} work on exoplanet
host star candidates.  These studies 
perform a careful (manual; expert) analysis using extremely high-resolution, high S/N spectra, and make use of \project{Hipparcos}
parallaxes where possible.  Most of the stars in these samples are main-sequence
or sub-giant stars.  Therefore, these works constitute an excellent comparison
to evaluate the accuracy of our results on the main-sequence and sub-giant branch.

We compare the stellar parameters ($\teff$, $\logg$, [Fe/H]) of stars that
overlap these literature studies in Figure~\ref{fig:gold-standard-comparison}.
We only include stars with $\chi_r^2 < 3$ and $S/N > 10$~pixel$^{-1}$, although the latter
cut removed only a few stars because the average S/N in the \rave\ spectra for
these stars is relatively high ($\gtrsim{}50$~pixel$^{-1}$). When considering the relative 
information content available in \rave\ (945~pixels in the near infrared with
$\mathcal{R} \approx 7{,}500$) compared to these literature studies that use
\project{Hipparcos} parallaxes where possible, and base their inferences on
spectra with resolving power $\mathcal{R}$ between 40,000 to 110,000, and
S/N ratios exceeding 150, we consider the agreement to be very satisfactory.
Indeed, given the metallicity precision available in the \unrave\ catalog, 
these results will likely be useful for future studies based on exoplanet
host star properties (e.g., \project{TESS}). At present, however, there are 
just $\approx$30 stars in \rave\ that overlap with the compilations of 
exoplanet host star properties listed at \texttt{exoplanets.org} and 
\texttt{exoplanets.eu}.

\subsubsection{Comparison with the \ges\ survey}
\label{sec:validation-ges}

There are 142 stars that overlap between \rave\ and the fourth internal
data release of the \ges\ survey. These are a mix of main-sequence, 
sub-giant and red giant branch stars.  About half (67) of the sample 
were acquired with the \acronym{UVES} instrument, the other with the \acronym{GIRAFFE} spectrograph, and
the S/N of the \ges\ spectra peaks at $\approx140$~pixel$^{-1}$.  Despite 
most of these stars having
relatively low S/N ratios in \rave\ ($\approx 25$~pixel$^{-1}$), there is good 
agreement in with \ges\ and the \unrave\ stellar parameters (Figure~\ref{fig:ges-stellar-parameters}).  
The RMS in effective temperature, surface gravity and metallicity is
233~K, 0.37~dex, and 0.17~dex, respectively.

Based on this comparison, we find no evidence for a systematic offset 
between stars on the main-sequence and those on the giant branch. Specifically,
we do not find any dwarf stars that are misclassified as sub-giants, as we
found for some metal-poor stars in Section~\ref{sec:validation-gold-standards}.
In addition we find no evidence for a systematic offset in metallicity (or
individual abundances) between giant stars and main-sequence stars.  This is
a crucial observation, as the metallicities for stars in our main-sequence
training set have a principally different source to those on the giant branch.
We cannot make these same inferences based on other surveys, like \apogee,
because (1) \apogee\ stars formed part of the training set, and (2) they do
not include main-sequence stars.  Even if we found good agreement between
\epic\ and \apogee\ metallicities, this would not be informative, because
\apogee\ is the source of metallicity for many stars on the giant branch in
the \epic\ sample. Therefore, although this is a qualitative comparison only,
it is reassuring that there is no obvious systematic difference between the
metallicities of main-sequence and giant branch stars.

The metallicity agreement between this work and \ges\ extends down to low
metallicity, near $[{\rm Fe/H}] \approx -1.5$.  The scatter increases for
the four stars in the overlap sample with [Fe/H] $< -1$, however these
spectra have low S/N ratios, which is reflected in the larger uncertainties
reported for these metallicities. 
% TODO: Check if that is still the case after we examine errors from repeat
% 		obs.

The fourth internal data release of the \ges\ includes detailed chemical
abundances of up to 45~species ($\approx32$ elements at different ionization
stages).  This provides us with an independent validation for our detailed
abundances on the giant branch.  These comparisons are shown for six elemental
abundances in Figure~\ref{fig:ges-abundances}, where markers are colored by the
S/N of the \rave\ spectrum.  The number of stars available in each abundance
comparison varies due to what is available in the \ges\ data release, which
is itself a function of the instrument used, the spectral type, and other
factors.  The absolute bias for individual elements varies from as low as
0.01~dex ([Al/H]) to as high as 0.22~dex ([Si/H]), where we over-estimate 
[Si/H] abundances relative to the \ges\ survey.  The large bias in [Si/H]
is likely a consequence of an offset between [Si/H] abundances in \ges\ and
\apogee, the source of our training set for giant star abundances.  
The RMS deviation in each
label is small for stars with [X/H] $> -0.5$, before increasing at lower
metallicities.  If we consider all stars, the smallest abundance RMS we see
with respect to \ges\ is 0.18~dex for [Ca/H].  The increasing RMS at low 
metallicity is likely a consequence of multiple factors, namely: inaccurate
abundance labels for metal-poor stars (Section~\ref{sec:the-training-set});
only weak, blended lines being available in \rave, which cease to be visible
in hot and/or metal-poor stars; and to a lesser extent, low S/N ratios for 
those particular stars being compared.  Unfortunately not all of these
factors are represented by the quoted errors in each label.  For these reasons, 
although it affects only a small number of stars, we recommend caution when 
using individual abundances for very metal-poor giant stars in our sample.

\subsection{Astrophysical validation}
\label{sec:astrophysical-validation}

\subsubsection{Globular clusters}
\label{sec:globular-cluster-validation}

After verifying that our atmospheric parameters and abundances are comparable with other
high-resolution studies, here we verify that the abundances we find
are consistent with expectations from astrophysics.  One example is globular clusters, 
which show anti-correlations in light element abundances 
\citep[e.g.,][and references therein]{Norris_Da_Costa_1995,Carretta_2009}.  In the \rave\
survey, \cite{Anguiano_2015} identified 70 stars with positions and radial velocities that are
consistent with being members of globular clusters: 49 stars belonging to NGC~5139
($\omega$~Centauri), 11 members of the retrograde globular cluster NGC~3201, and 10 
members of NGC~362. In addition to these systems, 12 stars thought to belong to NGC~1851
were found in the \rave\ data by \cite{Kunder_2014}, and a further 10 stars in NGC~6752.  
We refer the reader to those studies for details regarding the membership selection.

In Figure~\ref{fig:globular-cluster-HRD} we show the effective temperature
$\teff$ and surface gravity $\logg$ for pre-defined members of globular clusters in \rave.
The top panels indicate measurements made in this work, and for comparison purposes we have included
the results from the fourth \rave\ data release.  We show representative \project{PARSEC} isochrones 
\citep{Bressan_2012} in all panels, where the isochrone ages and metallicities are adopted from
\citet{Kunder_2016,Marin-Franch_2009}, and the \citet[][accessed 6 September 2016]{Harris_1996} 
online catalog of globular cluster properties.  The globular cluster with the most number of
\rave\ members is NGC~5139 ($\omega$-Centauri), where we find a significant metallicity spread
consistent with high-resolution studies \citep{Marino_2011,Carretta_2009,Carretta_2013}.  We
find the mean metallicity of prescribed $\omega$-Centauri members to be $[{\rm Fe/H}] = -0.85$.
However, it is clear that the membership criteria could be improved with our
revised metallicities and detailed chemical abundances. 
Indeed, this demonstrates that our abundance labels can be used to identify
globular cluster members that are now tidally disrupted \citep{Anguiano_2016,Kuzma_2016,Navin_2016},
even for stars with low S/N ratios.

\subsubsection{Open clusters}
\label{sec:open-cluster-validation}

The \rave\ survey observed $\sim$160 highly probable members of several open clusters:
78 stars in the Pleiades, 26 in Hyades, 30 stars in the solar-metallicity cluster
M67 and 13 stars in IC4651. Details on the open cluster membership and selection can 
be found in \cite{Kunder_2016}.  We show the effective temperature $\teff$ and surface
gravity $\logg$ for these stars in Figure \ref{fig:open-cluster-HRD}.  The isochrones
are sourced from \citet{Bressan_2012}, with cluster properties adopted from 
\cite{Kunder_2016} and \cite{Kharchenko_2013}.

We find good agreement between our atmospheric parameters (top panels) and the 
isochrones shown.  The position of the red clump in IC4651 and M67 are perfectly matched
to the isochrone, and the Hyades main-sequence is in good agreement down to 
$T_{\rm eff} \approx 4000$~K.  Similarly, we find consistency with the literature and
our metallicity scale.  We find the mean metallicity of M67 stars to be 
$[{\rm Fe/H}] = -0.02 \pm 0.17$~dex (30 stars), in excellent agreement with the
expected $[{\rm Fe/H}] = 0.00$ value (not accounting for atomic diffusion).
We further find the Hyades mean metallicity to be $[{\rm Fe/H}] = 0.07 \pm 0.09$~dex,
consistent with \citet{Paulson_2003}: $[{\rm Fe/H}] \approx 0.13$.
For IC4651 from 13 stars we find a mean $[{\rm Fe/H}] = 0.15 \pm 0.12$~dex, matching the
high-resolution, high S/N study of \citet{Pasquini_2004}, where they find 
$[{\rm Fe/H}] = 0.10 \pm 0.03$~dex. Finally, from 78 stars we find the mean
metallicity of the Pleiades to be $[{\rm Fe/H}] = -0.02 \pm 0.12$~dex, in agreement
with the $[{\rm Fe/H}] = -0.034 \pm 0.024$~dex measurement reported by 
\citet{Friel_Boesgaard_1990}.

The most pronounced discrepancy is seen in the main-sequence of the Pleiades.
Like \citet{Kordopatis_2013}, we find lower $\logg$ values than what the isochrone
shown predicts.  \stub{Hawkins: use correct isochrone from Bouy to see if this issue is resolved.}
%  Firstly,
%we note that theoretical isochrones often struggle to match the observed
%color-magnitude diagrams for some young clusters \citep[e.g., the Pleiades, see][]{Bouy_2015}.

\section{Discussion}
\label{sec:discussion}

We have derived atmospheric parameters ($\teff$, $\logg$, [Fe/H]) for all \rave\ 
spectra, as well as detailed abundances for red giant branch stars, based on an
independent re-analysis of \rave\ spectra.  Our catalog provides an excellent
complement to the current \rave\ data release \citep{Kunder_2016}.  
When combined with the \tgas\ sample, these results may amount to a powerful
compendium for chemo-dynamic studies of the Milky Way.  However, our analysis has caveats.
Inferences based on these results should recognize those caveats, and acknowledge that 
these results are subject to our explicit assumptions, some of which are provably
incorrect.

For practical purposes we adopted separate models: one for the giant branch and one
for the main-sequence.  A third model was used to derive relative weights for which
results to use.  The relative weighting we have used does not have any formal
interpretation as a likelihood or belief (in any sense): it was introduced for
practical reasons to identify systematic errors and combine results for multiple
models.  Because the relative weights have no formal interpretation, it is reasonable
to consider this method is as \emph{ad hoc} as any other approach.  Therefore, the
relative weighting has no warranty to be (formally) correct, and may introduce 
inconsistencies or systematic errors rather than
minimizing them.  Evidence of this is already present in our comparisons to
`gold standard' works \citep{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005},
where metal-poor dwarf stars appear as metal-poor sub-giants in our combined results,
but have reasonable labels (by any available statistic) as dwarfs in the main-sequence
model.  Clearly, the relative distance from the joint model to the giant model is 
slightly favoured (over the main-sequence model) for metal-poor dwarf stars. In this
sense, our approach towards combining labels from multiple models is sub-optimal.

If we only consider the results from individual models, there are a number of cautionary
remarks that stem from the construction of the training set.  The labels for red giant 
branch stars primarily come from \apogee, where previous successes with \thecannon\
have demonstrated that \apogee\ labels based on high S/N data can be of high fidelity
\citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}.
However, the lack of metal-poor stars in the \apogee/\rave\ overlap sample produced
a tapering-off in the test set --- where \emph{no} stars had reliably reported metallicities
below that of the training set --- which forced us to construct a heterogeneous training
set.  The metal-poor stars included in this sample are from high-resolution studies
\citep{Fulbright_2010,Ruchti_2011}, but it is unknown if the stellar parameters are
of high fidelity (e.g., we have limited quality metrics available), and there is no
guarantee that the stellar parameters \emph{or} abundances are on the same scale as
\apogee\ \citep[and good reasons to believe they will not be; see][]{Smiljanic_2014}.

If the metallicities of metal-poor giant stars are on the same scale as the \apogee\
abundances, there is a larger issue in verifying that the main-sequence metallicities
and giant branch metallicities are on the same scale.  The training set for the
main-sequence stars includes metallicities from a variety of sources --- including
\lamost, and the fourth \rave\ data release.  Even on expectation value, there is 
no straightforward manner to ensure that these two models produce metallicities on
the same scale.  We see no systematic offset in metallicities of dwarf and giant
stars that overlap between \rave\ and the \ges\ survey, suggesting that if there is
a systematic offset, it must be small. Nevertheless, these are only verification checks based on $<1$\% of the
data, and there is currently insufficient data for us to prove both models are on
the same abundance scale.

For some of the most metal-poor giant stars in \rave\, we \emph{know} the abundances 
are not on the same scale as \apogee, because we were forced to adopt abundances for 
specific elements when they were unavailable.  In that sense, our assumption that the
training set contains accurate labels is clearly untrue (for \emph{at least} some stars).
Although we sought to adopt the mean
level of Galactic chemical enrichment at the overall metallicity, this is not a
representative abundance. Even if that is the \emph{mean} enrichment at that Galactic
metallicity, there is no requirement for zero abundance spread.
More fundamentally, we are incorrectly assuming that the element \emph{must} be 
detectable in the photosphere of the star.  There may be no transition that is 
detectable in that star, even with zero noise, because it is too weak to have any 
effect on the spectrum.  In the most optimistic case, this can be considered to be
forcing the model to make use of correlated information between abundances.  In a
more representative (pessimistic) case, we are simply invoking what all abundances
should be at low metallicity.

This choice is reflected in the abundances of the test set.  While we do recover
trustworthy metallicities for ultra metal-poor ($[{\rm Fe/H}] \lesssim -4$) stars like 
CD-38~245, the individual abundances for all extremely metal-poor stars aggregate
(in [X/Fe] space; Figure~\ref{fig:gce}) at the assumed abundances for the metal-poor
stars in our sample.  Thus, while the overall metallicities appear reliable, the 
individual abundances for extremely metal-poor stars in the test set cannot be
considered trustworthy in any sense.  For this reason we have updated the electronic
catalog to discard these results, and included a quality control flag to indicate
the result was discarded (see catalog \texttt{README} file for specifics related to
the data structure and content).

In Section~\ref{sec:method} we assumed that any fibre- or time-dependent variations
in the \rave\ spectra are negligible.  This is provably incorrect. Indeed,
\citet{Kordopatis_2013} note that the effective resolution of \rave\
spectra varies from $6{,}500 < \mathcal{R} < 8{,}500$, and that the 
effective resolution is a function of temperature variations, fibre-to-fibre
variations, and thus position on the CCD \citep{Steinmetz_2006}.
 For this reason we ought to expect our derived stellar
parameters or abundances to be correlated with the fibre number, or over
time.  If significant, the trend could produce systematically offset
stellar abundances solely due to the fibre number.  \citet{Kordopatis_2013}
conclude that resolution-based effects on the \rave\ stellar parameters
should be a second-order effect. We have not seen evidence of these 
resolution-based correlations in our results, however, we have only
performed cursory (non-exhaustive) experiments to investigate this issue.

\stub{assumptions that are probably wrong: fringing in high S/N spectra produces issues with the normalization and thus the spectrum. KCF to comment more on this}

The fibre fringing is another example of how test spectra that are
unusual from the training set can be projected as peculiar artefacts in the
label space.  In other words, unusual spectra can appear as `clumps' in
regions of parameter space that we could consider as being normal 
(e.g., solar-type stars).  We addressed this issue for the main-sequence
and giant models by using a joint model to calculate relative weights.
However, spectra that are unusual from the joint model training set can
still project as systematic artefacts.

Indeed, there are two known artefacts in our data that are relevant to this discusssion. The
first is a small over-density at the base of the giant branch, which is
likely a consequence of joining the 9-label and 3-label models.  The
second has an astrophysical origin: there are no hot stars ($\teff > 8000$~K)
present in our training set, yet there are many in the \rave\ survey.
However, the \rave\ pre-processing pipeline (\texttt{SPARV}) performs
template matching against grids of cool \emph{and} hot stars, and therefore
we can use those results to identify hot stars. In Figure \ref{fig:hot-stars} 
we show our derived effective temperatures $\teff$ and surface gravities $\logg$,
where each hexagonal bin is colored by the \emph{maximum} temperature
reported by \texttt{SPARV} for any star in that bin.  Hot stars dominate
at just above the turn-off point in our results.

>>>>

\begin{itemize}
\item We assume that any fibre- and time-dependent variations in spectral
resolution in the \rave\ spectra are negligible.
\item The \rave\ noise variances are approximately correct, independent between
pixels, and normally distributed.
\item We assume that the normalization procedure employed by the \rave\ pipeline
is invariant with respect to the labels we seek to measure (e.g., $\teff$, $\logg$,
or [Fe/H]), and invariant with respect to the S/N ratio.  In other words, we assume
that the normalization procedure does not produce different results for high S/N
spectra compared to low S/N spectra, nor does the normalization procedure vary 
non-linearly with respect to stellar parameters (e.g., [Fe/H]).
\item We assume that stars with similar labels ($\teff$, $\logg$, and abundances)
have similar spectra.
\item A stellar spectrum is a smooth function of the label values for that star,
and we assume that the function is smooth enough within a sub-space of the labels
(e.g., the giant branch or the main-sequence) that it can be reasonably approximated 
with a low-order polynomial in label space.
\item The training set (Section~\ref{sec:the-training-set}) has mean accurate labels
for most, but not all stars. That is to say that we do not assume that \emph{every} 
label in the training is accurate -- we
can afford to have a small fraction of inaccurate labels (e.g., a few obvious 
misclassifications in the training set are affordable).
\item We assume that the training data are similar (in spectra) to the test data 
where they overlap in label space, and we assume that the training data spans enough
of the label space to capture the variation in the test-set spectra.
\end{itemize}

@@

We
